### 2025-06-10 correlation.qmd
20250610 - cor to cov

```{r}
player_stats_seasonal %>% 
  dplyr::select(height, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

@@

$$
\text{Cov}(x, y) = \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})
$$ {#eq-covariance}

@@

>>>>

```{r}
player_stats_seasonal %>% 
  dplyr::select(height, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        CV = ~ sd(., na.rm = TRUE) / mean(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

@@

$$
\begin{aligned}
  \text{Cov}(x, y) &= \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})\\
  &= r_{xy} \cdot \sigma_x \cdot \sigma_y
\end{aligned}
$$ {#eq-covariance}

where $n$ is the sample size, $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $x_i$ is the ith observation on the [predictor variable](#sec-correlationalStudy), $y_i$ is the ith observation of the [outcome variable](#sec-correlationalStudy), $\bar{x}$ is the mean of the [predictor variable](#sec-correlationalStudy), $\bar{y}$ is the mean of the [outcome variable](#sec-correlationalStudy), $r_{xy}$ is the correlation between the [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variables, $\sigma_x$ is the standard deviation of the [predictor variable](#sec-correlationalStudy), and $\sigma_y$ is the standard deviation of the [outcome variable](#sec-correlationalStudy).

@@

We can also convert a correlation back to covariance using the formula in @eq-covariance.

```{r}
cor_value * sd(complete_cases$height, na.rm = TRUE) * sd(complete_cases$fantasyPoints, na.rm = TRUE)
```
